---
title: "Data overlap: Enriched heatmaps"
author: "Deepak Tanwar"
date: "<b>Created on:</b> 2020-10-09 <br> <b>Updated on:</b> `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: tango
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    keep_md: no
    number_sections: no
    fig_width: 8
    fig_height: 8
    fig_caption: true
    df_print: paged
    code_folding: show
  fontsize: 12pt
  geometry: margin=1in
  documentclass: article
# bibliography: references.bib
link-citations: yes
---

# Library
```{r enriched-heatmaps-data-overlap-1, warning = F, message = F}
library(EnrichedHeatmap)
library(circlize)
library(data.table)
library(tidyverse)
library(plgINS)
library(BSgenome.Mmusculus.UCSC.mm10)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(csaw)
library(gUtils)
library(report)
library(plyr)
library(viridis)
library(RColorBrewer)
```

## Heatmap options
```{r enriched-heatmaps-data-overlap-2 }
ht_opt$TITLE_PADDING <- unit(1, "mm")
ht_opt$legend_gap <- unit(3, "mm")
ht_opt$legend_grid_height <- unit(2, "mm")
ht_opt$legend_grid_width <- unit(2, "mm")
ht_opt$HEATMAP_LEGEND_PADDING <- unit(1, "mm")
```


# Function to run in parallel
```{r enriched-heatmaps-data-overlap-3, echo = F}
mylapply <- function(...) {
  if (require(parallel) && .Platform$OS.type == "unix") {
    mclapply(..., mc.preschedule = F)
  }
  else {
    lapply(...)
  }
}
```

# ATAC differential analysis
```{r enriched-heatmaps-data-overlap-4 }
load("input/atac_diff_chromatin_accessibility.RData")

res <- data.frame(rowData(data), stringsAsFactors = F, check.names = F)
colnames(res) <- gsub(pattern = "Peak-", replacement = "", x = colnames(res))
res_p <- res[res$`diffAccessibility-qvalue` <= 0.05, ]
res_lp <- res[res$`diffAccessibility-qvalue` <= 0.05 & abs(res$`diffAccessibility-logFC`) > 1, ]

tss_mm10 <- gr.mid(GRanges(res_lp))
names(tss_mm10) <- tss_mm10$Name

tss_5k <- promoters(tss_mm10, upstream = 1000, downstream = 1000)
names(tss_5k) <- tss_5k$Name
```

# Data overlaps
```{r enriched-heatmaps-data-overlap-5 }
load("input/overlap_tables.RData")
wo <- ldply(sp_df_s1, data.frame)[, -1]
wo <- inner_join(res_lp, wo)
wi <- ldply(sp_df_s2, data.frame)[, -1]
wi <- inner_join(res_lp, wi)
```


# Normalized matrix
```{r enriched-heatmaps-data-overlap-6 }
nm_atac <- "./input/mat_atac.RData"
nm_rna_counts <- "./input/mat_rna_counts.RData"
nm_rna_logFC <- "./input/mat_rna_counts_logFC.RData"
nm_chip <- "./input/mat_chip.RData"
nm_bs <- "./input/mat_bs.RData"

load(nm_atac)
load(nm_rna_counts)
load(nm_rna_logFC)
load(nm_chip)
load(nm_bs)
```

# Heatmaps functions

## Color function
```{r enriched-heatmaps-data-overlap-7 }
generate_diff_color_fun <- function(x, rows) {
  q <- quantile(x[rows, ], c(0.5, .95))
  max_q <- max(abs(q))
  colorRamp2(c(-max_q, 0, max_q), c("#3794bf", "#FFFFFF", "#df8640"))
}

generate_color_fun <- function(x, rows, col = "red") {
  q <- NULL
  if (is.list(x)) {
    a <- lapply(x, function(y) quantile(data.matrix(y)[rows, ], c(0.05, 0.95)))
    min <- NULL
    max <- NULL
    for (i in 1:length(a)) {
      min <- c(min, a[[i]][1])
      max <- c(max, a[[i]][2])
    }
    q <- c(min(min), max(max))
  } else {
    q <- quantile(x[rows, ], c(0.05, 0.95))
  }
  max_q <- max(abs(q))
  col <- colorRamp2(c(0, max_q), c("#FFFFFF", col))
  return(col)
}
```

## Function for `at`
```{r enriched-heatmaps-data-overlap-8 }
get_at <- function(mm) {
  l <- length(mm@levels)
  mm@colors <- c(mm@colors[1], mm@colors[median(1:l)], mm@colors[l])
  mm@levels <- c(mm@levels[1], mm@levels[median(1:l)], mm@levels[l])
  return(mm)
}
```

## Fold change color for heatmap (from Pierre-Luc)
```{r enriched-heatmaps-data-overlap-9}
sqrtCols <- function(x, cols = NULL) {
  if (is.null(cols)) {
    cols <- rev(brewer.pal(n = 11, "RdBu"))
  }
  if (is.function(cols)) {
    cols <- cols(11)
  }
  if (length(cols) != 11) {
    stop("`cols` should contain 11 colors.")
  }
  bb <- c(
    -seq(from = sqrt(abs(min(x, na.rm = TRUE))), to = 0, length.out = 6)^2,
    seq(from = 0, to = sqrt(max(x, na.rm = TRUE)), length.out = 6)[-1]^2
  )

  colorRamp2(bb, cols)
}
```


## Enriched Heatmap creating function
```{r enriched-heatmaps-data-overlap-10 }
make_EH <- function(data, name, color, column_title, anno_name, show_legend,
                    log2 = FALSE, bg_col, col_an, partition, axis_name, ht) {
  if (log2) {
    data <- log2(data + 1)
  }

  h <- EnrichedHeatmap(
    mat = data,
    name = name,
    col = color,
    column_title = column_title,
    top_annotation = HeatmapAnnotation(
      lines = anno_enriched(
        height = unit(1, "cm"),
        gp = gpar(
          lwd = 0.7,
          fontsize = 6,
          fontfamily = "Helvetica",
          col = col_an,
          lty = 1:length(unique(col_an))
        ),
        axis_param = list(
          side = "right",
          facing = "inside",
          gp = gpar(
            fontsize = 6,
            col = "black",
            lwd = 0.4
          )
        )
      ),
      annotation_name_gp = gpar(
        fontsize = 0,
        fontfamily = "Helvetica"
      ),
      annotation_legend_param = list(
        title_gp = gpar(
          fontsize = 6,
          fontface = "bold",
          fontfamily = "Helvetica"
        ),
        labels_gp = gpar(
          fontsize = 6,
          fontfamily = "Helvetica"
        )
      )
    ),
    border_gp = gpar(
      col = "black",
      lwd = 0.4
    ),
    column_title_gp = gpar(
      fontfamily = "Helvetica",
      fontsize = 6,

      fill = bg_col
    ),
    axis_name_gp = gpar(
      fontfamily = "Helvetica",
      fontsize = 6,
      col = "black",
      lwd = 0.4
    ),
    use_raster = FALSE,
    row_split = partition,
    row_gap = unit(0.5, "mm"),
    row_title = NULL,
    axis_name = axis_name,
    width = unit(1.8, "cm"),
    height = unit(ht, "cm"),
    show_heatmap_legend = show_legend,
    heatmap_legend_param = list(
      direction = "vertical",
      # at = get_at(m1 = mat_AS$Adult_NF[x,], m2 = mat_AS$PND15_NF[x,]),
      title = anno_name,
      legend_height = unit(0.8, "cm"),
      title_position = "topleft",
      title_gp = gpar(fontsize = 6, fontfamily = "Helvetica"),
      labels_gp = gpar(fontsize = 6, fontfamily = "Helvetica")
    )
  )
  return(h)
}
```


## EH function
```{r enriched-heatmaps-data-overlap-11 }
draw_EH <- function(x, k, name, ht, type, log = FALSE, outdir, rm) {
  set.seed(1100)

  names(k) <- x
  partition <- k

  nc <- length(unique(k))

  pdf_file <- NULL

  if (log) {
    if (rm == "meth") {
      pdf_file <- paste0("./output/", outdir, "/", name, "_no_meth_log2.pdf")
    } else {
      pdf_file <- paste0("./output/", outdir, "/", name, "_log2.pdf")
    }
  } else {
    if (rm == "meth") {
      pdf_file <- paste0("./output/", outdir, "/", name, "_no_meth.pdf")
    } else {
      pdf_file <- paste0("./output/", outdir, "/", name, ".pdf")
    }
  }


  col_an <- NULL
  if (length(unique(partition)) <= 8) {
    col_an <- brewer.pal(n = 8, name = "Dark2")
    names(col_an) <- unique(partition)

    col_an <- col_an[!is.na(names(col_an))]
  } else {
    n <- length(unique(partition))
    qual_col_pals <- brewer.pal.info[brewer.pal.info$category == "qual", ]
    col_an <- unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
    names(col_an) <- unique(partition)
    col_an <- col_an[!is.na(names(col_an))]
    # pie(rep(1,n), col=sample(col_vector, n))
  }

  hm_an <- HeatmapAnnotation(
    Categories = partition,
    which = "row",
    annotation_name_gp = gpar(fontsize = 0),
    annotation_legend_param = list(
      title_gp = gpar(
        fontsize = 6,
        fontface = "bold",
        fontfamily = "Helvetica"
      ),
      labels_gp = gpar(
        fontsize = 6,
        fontfamily = "Helvetica"
      )
    ),
    col = list(Categories = col_an),
    simple_anno_size = unit(3, "mm")
  )

  axis_name <- c("-1kb", "Mid", "1kb")

  mat_as <- lapply(mat_AS, function(y) {
    if (log) {
      return(log2(y[x, ] + 1))
    } else {
      return(y[x, ])
    }
  })

  col_atac <- generate_color_fun(x = mat_as, col = "red", rows = x)

  bg_col <- brewer.pal(n = 5, name = "Set2")

  # ATAC-Seq
  atac_PND15 <- make_EH(
    data = mat_AS$PND15_NF[x, ],
    name = "PND15",
    column_title = "PND15 ATAC",
    color = col_atac,
    anno_name = ifelse(log, expression(bold(ATAC ~ log[2] ~ "RPKM")), expression(bold(ATAC ~ "RPKM"))),
    show_legend = FALSE, log2 = log, bg_col = bg_col[1], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  atac_Adult <- make_EH(
    data = mat_AS$Adult_NF[x, ],
    name = "Adult",
    column_title = "Adult ATAC",
    color = col_atac,
    anno_name = ifelse(log, expression(bold(ATAC ~ log[2] ~ "RPKM")), expression(bold(ATAC ~ "RPKM"))),
    show_legend = TRUE, log2 = log, bg_col = bg_col[1], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  hm_atac <- atac_PND15 + atac_Adult

  hm_atac@ht_list$PND15@matrix_color_mapping <- get_at(mm = hm_atac@ht_list$PND15@matrix_color_mapping)
  hm_atac@ht_list$Adult@matrix_color_mapping <- get_at(mm = hm_atac@ht_list$Adult@matrix_color_mapping)


  chip_col_dat <- function(z) {
    if (log) {
      return(log2(z[x, ] + 1))
    } else {
      return(z[x, ])
    }
  }

  chip_PNW8_H3K4me3 <- make_EH(
    data = mat_CS$PNW8_H3K4me3_SRX332343[x, ],
    name = "PNW8 H3K4me3",
    column_title = "PNW8 H3K4me3",
    color = generate_color_fun(x = chip_col_dat(z = mat_CS$PNW8_H3K4me3_SRX332343[x, ]), rows = x, col = "blue"),
    anno_name = ifelse(log, expression(bold(H3K4me3 ~ log[2] ~ "RPKM")), expression(bold(H3K4me3 ~ "RPKM"))),
    show_legend = TRUE, log2 = log, bg_col = bg_col[2], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  chip_PNW8_H3K27ac <- make_EH(
    data = mat_CS$PNW8_H3K27ac_SRX332351[x, ],
    name = "PNW8 H3K27ac",
    column_title = "PNW8 H3K27ac",
    color = generate_color_fun(x = chip_col_dat(z = mat_CS$PNW8_H3K27ac_SRX332351[x, ]), rows = x, col = "blue"),
    anno_name = ifelse(log, expression(bold(H3K27ac ~ log[2] ~ "RPKM")), expression(bold(H3K27ac ~ "RPKM"))),
    show_legend = TRUE, log2 = log, bg_col = bg_col[2], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  chip_PNW8_H3K27me3 <- make_EH(
    data = mat_CS$PNW8_H3K27me3_SRX332346[x, ],
    name = "PNW8 H3K27me3",
    column_title = "PNW8 H3K27me3",
    color = generate_color_fun(x = chip_col_dat(z = mat_CS$PNW8_H3K27me3_SRX332346[x, ]), rows = x, col = "blue"),
    anno_name = ifelse(log, expression(bold(H3K27me3 ~ log[2] ~ "RPKM")), expression(bold(H3K27me3 ~ "RPKM"))),
    show_legend = TRUE, log2 = log, bg_col = bg_col[2], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )


  hm_chip <- chip_PNW8_H3K4me3 + chip_PNW8_H3K27ac + chip_PNW8_H3K27me3

  hm_chip@ht_list$`PNW8 H3K4me3`@matrix_color_mapping <- get_at(hm_chip@ht_list$`PNW8 H3K4me3`@matrix_color_mapping)
  hm_chip@ht_list$`PNW8 H3K27ac`@matrix_color_mapping <- get_at(hm_chip@ht_list$`PNW8 H3K27ac`@matrix_color_mapping)
  hm_chip@ht_list$`PNW8 H3K27me3`@matrix_color_mapping <- get_at(hm_chip@ht_list$`PNW8 H3K27me3`@matrix_color_mapping)


  # Methylation color
  meth_col_fun <- colorRamp2(c(0, 0.5, 1), c("red", "white", "blue"))


  # BSseq

  bs_PND7 <- make_EH(
    data = mat_BS$PND7_SRX749887[x, ],
    name = "PND7 BS",
    column_title = "PND7 BS",
    color = meth_col_fun,
    anno_name = expression(bold(DNAme)),
    show_legend = TRUE, log2 = FALSE, bg_col = bg_col[4], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  bs_PND14 <- make_EH(
    data = mat_BS$PND14_SRX749892[x, ],
    name = "PND14 BS",
    column_title = "PND14 BS",
    color = meth_col_fun,
    anno_name = expression(bold(DNAme)),
    show_legend = FALSE, log2 = FALSE, bg_col = bg_col[4], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  bs_PNW8 <- make_EH(
    data = mat_BS$PNW8[x, ],
    name = "PNW8 BS",
    column_title = "PNW8 BS",
    color = meth_col_fun,
    anno_name = expression(bold(DNAme)),
    show_legend = FALSE, log2 = FALSE, bg_col = bg_col[4], col_an = col_an,
    partition = partition, axis_name = axis_name, ht = ht
  )

  hm_bs <- bs_PND7 + bs_PND14 + bs_PNW8

  cols_rna_logFC <- brewer.pal(n = 3, name = "Dark2")[1:2]
  cols_rna_logFC <- c(`PND8 vs PND15` = cols_rna_logFC[1], `PND14 vs PNW8` = cols_rna_logFC[2])

  ta_logFC <- HeatmapAnnotation(
    Comparison = factor(c("PND8 vs PND15", "PND14 vs PNW8"), levels = c("PND8 vs PND15", "PND14 vs PNW8")),
    col = list(Comparison = cols_rna_logFC),
    simple_anno_size = unit(3, "mm"),
    show_annotation_name = TRUE,
    height = unit(3, "mm"),
    annotation_name_side = "right",
    annotation_name_rot = 0,
    annotation_name_gp = gpar(fontsize = 0),
    annotation_legend_param = list(
      title_gp = gpar(fontsize = 6, fontface = "bold", fontfamily = "Helvetica"),
      labels_gp = gpar(fontsize = 6, fontfamily = "Helvetica")
    )
  )

  hm_rna_counts_logFC <- Heatmap(
    matrix = data.matrix(mat_RNA_counts_logFC[x, ]),
    col = sqrtCols(mat_RNA_counts_logFC[x, ]),
    name = "RNA logFC",
    na_col = "grey",
    show_row_names = FALSE,
    show_column_names = FALSE,
    cluster_rows = F,
    cluster_columns = F,
    width = unit(0.5, "cm"),
    height = unit(ht, "cm"),
    use_raster = FALSE,
    row_split = partition,
    row_title = NULL,
    top_annotation = ta_logFC,
    heatmap_legend_param = list(
      direction = "vertical",
      # at = c(-4, -2, 0, 2, 4),
      title = expression(bold(~ Log[2] ~ "fold-change")),
      legend_height = unit(0.8, "cm"),
      title_position = "topleft",
      title_gp = gpar(fontsize = 6, fontfamily = "Helvetica"),
      labels_gp = gpar(fontsize = 6, fontfamily = "Helvetica")
    )
  )

  hm_rna_counts_logFC@matrix_color_mapping <- get_at(mm = hm_rna_counts_logFC@matrix_color_mapping)

  if (type == "wo") {
    if (all(is.na(mat_RNA_counts[x, ]))) {
      ht_list <- hm_an + hm_atac
    } else if (name == "distal" | name == "distal_accUp" | name == "distal_accDown" | name == "proximal_inactive") {
      ht_list <- hm_an + hm_atac
    } else {
      ht_list <- hm_an + hm_atac + hm_rna_counts_logFC
    }
  } else {
    if (all(is.na(mat_RNA_counts[x, ]))) {
      if (rm == "meth") {
        ht_list <- hm_an + hm_atac + hm_chip
      } else {
        ht_list <- hm_an + hm_atac + hm_chip + hm_bs
      }
    } else if (name == "distal" | name == "distal_accUp" | name == "distal_accDown" | name == "proximal_inactive") {
      if (rm == "meth") {
        ht_list <- hm_an + hm_atac + hm_chip
      } else {
        ht_list <- hm_an + hm_atac + hm_chip + hm_bs
      }
    } else {
      if (rm == "meth") {
        ht_list <- hm_an + hm_atac + hm_chip + hm_rna_counts_logFC
      } else {
        ht_list <- hm_an + hm_atac + hm_chip + hm_bs + hm_rna_counts_logFC
      }
    }
  }


  ht_opt$TITLE_PADDING <- unit(1, "mm")
  ht_opt$legend_gap <- unit(3, "mm")
  ht_opt$legend_grid_height <- unit(2, "mm")
  ht_opt$legend_grid_width <- unit(2, "mm")
  ht_opt$HEATMAP_LEGEND_PADDING <- unit(1, "mm")

  pdf(file = pdf_file, width = 11, height = 8.5)

  grid.newpage()
  pushViewport(viewport(gp = gpar(lwd = 0.5)))
  draw(
    ht_list,
    ht_gap = unit(1, "mm"),
    merge_legends = TRUE,
    newpage = FALSE
  )
  popViewport()

  dev.off()

  png_file <- gsub(pattern = "pdf", replacement = "png", x = pdf_file)

  png(filename = png_file, width = 11, height = 8.5, units = "in", res = 330)
  grid.newpage()
  pushViewport(viewport(gp = gpar(lwd = 0.5)))
  draw(
    ht_list,
    ht_gap = unit(1, "mm"),
    merge_legends = TRUE,
    newpage = FALSE
  )
  popViewport()

  dev.off()

  invisible(gc())
}
```


# Enriched heatmaps

## Replace names
```{r enriched-heatmaps-data-overlap-12 }
change_names <- function(n) {
  n <- gsub(pattern = "active", replacement = "Active", x = n)
  n <- gsub(pattern = "inActive", replacement = "Inactive", x = n)
  n <- gsub(pattern = "accUp", replacement = "More Acc", x = n)
  n <- gsub(pattern = "accDown", replacement = "Less Acc", x = n)
  n <- gsub(pattern = "rnaUp", replacement = "More Expr", x = n)
  n <- gsub(pattern = "rnaDown", replacement = "Less Expr", x = n)
  n <- gsub(pattern = "\\.", replacement = " & ", x = n)
  n <- gsub(pattern = "More Acc & More Expr", replacement = "Category 1", x = n)
  n <- gsub(pattern = "More Acc & Less Expr", replacement = "Category 2", x = n)
  n <- gsub(pattern = "Less Acc & Less Expr", replacement = "Category 3", x = n)
  n <- gsub(pattern = "Less Acc & More Expr", replacement = "Category 4", x = n)
  n <- gsub(pattern = "Acc", replacement = "acc.", x = n)
  n <- gsub(pattern = "& H3K27me3", replacement = "(H3K27me3)", x = n)
  n <- gsub(pattern = "& H3K4me3", replacement = "(H3K4me3)", x = n)
  n <- gsub(pattern = "& H3K27ac", replacement = "(H3K27ac)", x = n)
  n <- gsub(pattern = "Less acc.$", replacement = "Less acc. (No H3 mark)", x = n)
  n <- gsub(pattern = "More acc.$", replacement = "More acc. (No H3 mark)", x = n)
  n <- gsub(pattern = "Category 1$", replacement = "Category 1 (No H3 mark)", x = n)
  n <- gsub(pattern = "Category 2$", replacement = "Category 2 (No H3 mark)", x = n)
  n <- gsub(pattern = "Category 3$", replacement = "Category 3 (No H3 mark)", x = n)
  n <- gsub(pattern = "Category 4$", replacement = "Category 4 (No H3 mark)", x = n)

  return(n)
}
```

## Without ChIP data

### Distal regions
```{r enriched-heatmaps-data-overlap-13 }
wo_distal <- wo[grep(pattern = "distal", x = wo$anno), ]
wo_k_distal <- gsub(pattern = "distal.", replacement = "", x = wo_distal$anno)
wo_k_distal <- change_names(n = wo_k_distal)

draw_EH(x = wo_distal$Name, k = wo_k_distal, name = "distal", ht = 6.5, type = "wo", log = FALSE, outdir = "without_ChIP", rm = "none")
draw_EH(x = wo_distal$Name, k = wo_k_distal, name = "distal", ht = 6.5, type = "wo", log = TRUE, outdir = "without_ChIP_log2", rm = "none")
```

### Proximal regions
```{r enriched-heatmaps-data-overlap-14 }
wo_pr <- wo[grep(pattern = "proximal", x = wo$anno), ]
wo_pr <- split(x = wo_pr, f = wo_pr$anno)
wo_pr <- wo_pr[lapply(wo_pr, nrow) > 10]
wo_proximal <- plyr::ldply(wo_pr, data.frame)[, -1]
wo_k_proximal <- gsub(pattern = "proximal.", replacement = "", x = wo_proximal$anno)
wo_k_proximal <- change_names(n = wo_k_proximal)

draw_EH(x = wo_proximal$Name, k = wo_k_proximal, name = "proximal", ht = 7, type = "wo", log = FALSE, outdir = "without_ChIP", rm = "none")
draw_EH(x = wo_proximal$Name, k = wo_k_proximal, name = "proximal", ht = 7, type = "wo", log = TRUE, outdir = "without_ChIP_log2", rm = "none")
```

### Proximal active regions
```{r enriched-heatmaps-data-overlap-15 }
wo_pr_ac <- wo[grep(pattern = "proximal.active", x = wo$anno), ]
wo_pr_ac <- split(x = wo_pr_ac, f = wo_pr_ac$anno)
wo_pr_ac <- wo_pr_ac[lapply(wo_pr_ac, nrow) > 10]
wo_proximal_active <- plyr::ldply(wo_pr_ac, data.frame)[, -1]
wo_k_ac <- gsub(pattern = "proximal.active.", replacement = "", x = wo_proximal_active$anno)
wo_k_ac <- change_names(n = wo_k_ac)
wo_k_ac <- trimws(gsub(pattern = "\\(No H3 mark)", replacement = "", x = wo_k_ac))
wo_k_ac <- factor(x = wo_k_ac, levels = c("Category 1", "Category 2", "Category 3", "Category 4"))


draw_EH(x = wo_proximal_active$Name, k = wo_k_ac, name = "proximal_active", ht = 5.2, type = "wo", log = FALSE, outdir = "without_ChIP", rm = "none")

draw_EH(x = wo_proximal_active$Name, k = wo_k_ac, name = "proximal_active", ht = 5.2, type = "wo", log = TRUE, outdir = "without_ChIP_log2", rm = "none")
```


### Proximal inactive regions
```{r enriched-heatmaps-data-overlap-16 }
wo_pr_inac <- wo[grep(pattern = "proximal.inactive", x = wo$anno), ]
wo_pr_inac <- split(x = wo_pr_inac, f = wo_pr_inac$anno)
wo_pr_inac <- wo_pr_inac[lapply(wo_pr_inac, nrow) > 10]
wo_proximal_inactive <- plyr::ldply(wo_pr_inac, data.frame)[, -1]
wo_k_inac <- gsub(pattern = "proximal.inactive.", replacement = "", x = wo_proximal_inactive$anno)
wo_k_inac <- change_names(n = wo_k_inac)
wo_k_inac <- trimws(gsub(pattern = "\\(No H3 mark)", replacement = "", x = wo_k_inac))

draw_EH(x = wo_proximal_inactive$Name, k = wo_k_inac, name = "proximal_inactive", ht = 5.2, type = "wo", log = FALSE, outdir = "without_ChIP", rm = "none")
draw_EH(x = wo_proximal_inactive$Name, k = wo_k_inac, name = "proximal_inactive", ht = 5.2, type = "wo", log = TRUE, outdir = "without_ChIP_log2", rm = "none")
```


## With ChIP data

### Distal regions
```{r enriched-heatmaps-data-overlap-17 }
wi_dis <- wi[grep(pattern = "distal", x = wi$anno), ]
wi_dis <- split(x = wi_dis, f = wi_dis$anno)
wi_dis <- wi_dis[lapply(wi_dis, nrow) > 10]
wi_distal <- plyr::ldply(wi_dis, data.frame)[, -1]
wi_k_distal <- gsub(pattern = "distal.", replacement = "", x = wi_distal$anno)
wi_k_distal <- change_names(n = wi_k_distal)

wi_k_distal <- factor(wi_k_distal, unique(wi_k_distal))

draw_EH(x = wi_distal$Name, k = wi_k_distal, name = "distal", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_distal$Name, k = wi_k_distal, name = "distal", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_distal$Name, k = wi_k_distal, name = "distal", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_distal$Name, k = wi_k_distal, name = "distal", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```

### Distal regions accUp
```{r enriched-heatmaps-data-overlap-18 }
wi_dis_acup <- wi[grep(pattern = "distal.accUp", x = wi$anno), ]
wi_dis_acup <- split(x = wi_dis_acup, f = wi_dis_acup$anno)
wi_dis_acup <- wi_dis_acup[lapply(wi_dis_acup, nrow) > 10]
wi_distal_acup <- plyr::ldply(wi_dis_acup, data.frame)[, -1]
wi_k_distal_acup <- gsub(pattern = "distal.", replacement = "", x = wi_distal_acup$anno)
wi_k_distal_acup <- change_names(n = wi_k_distal_acup)

wi_k_distal_acup <- factor(wi_k_distal_acup, unique(wi_k_distal_acup))

draw_EH(x = wi_distal_acup$Name, k = wi_k_distal_acup, name = "distal_accUp", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_distal_acup$Name, k = wi_k_distal_acup, name = "distal_accUp", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_distal_acup$Name, k = wi_k_distal_acup, name = "distal_accUp", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_distal_acup$Name, k = wi_k_distal_acup, name = "distal_accUp", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```

### Distal regions accDown
```{r enriched-heatmaps-data-overlap-19 }
wi_dis_acdn <- wi[grep(pattern = "distal.accDown", x = wi$anno), ]
wi_dis_acdn <- split(x = wi_dis_acdn, f = wi_dis_acdn$anno)
wi_dis_acdn <- wi_dis_acdn[lapply(wi_dis_acdn, nrow) > 10]
wi_distal_acdn <- plyr::ldply(wi_dis_acdn, data.frame)[, -1]
wi_k_distal_acdn <- gsub(pattern = "distal.", replacement = "", x = wi_distal_acdn$anno)
wi_k_distal_acdn <- change_names(wi_k_distal_acdn)

wi_k_distal_acdn <- factor(wi_k_distal_acdn, unique(wi_k_distal_acdn))

draw_EH(x = wi_distal_acdn$Name, k = wi_k_distal_acdn, name = "distal_accDown", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_distal_acdn$Name, k = wi_k_distal_acdn, name = "distal_accDown", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_distal_acdn$Name, k = wi_k_distal_acdn, name = "distal_accDown", ht = 12, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_distal_acdn$Name, k = wi_k_distal_acdn, name = "distal_accDown", ht = 12, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```


### Proximal regions
```{r enriched-heatmaps-data-overlap-20 }
wi_pr <- wi[grep(pattern = "proximal", x = wi$anno), ]
wi_pr <- split(x = wi_pr, f = wi_pr$anno)
wi_pr <- wi_pr[lapply(wi_pr, nrow) > 10]
wi_proximal <- plyr::ldply(wi_pr, data.frame)[, -1]
wi_k_proximal <- gsub(pattern = "proximal.", replacement = "", x = wi_proximal$anno)
wi_k_proximal <- change_names(wi_k_proximal)

wi_k_proximal <- factor(wi_k_proximal, unique(wi_k_proximal))

draw_EH(x = wi_proximal$Name, k = wi_k_proximal, name = "proximal", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_proximal$Name, k = wi_k_proximal, name = "proximal", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_proximal$Name, k = wi_k_proximal, name = "proximal", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_proximal$Name, k = wi_k_proximal, name = "proximal", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```

### Proximal active regions
```{r enriched-heatmaps-data-overlap-21 }
wi_pr_ac <- wi[grep(pattern = "proximal.active", x = wi$anno), ]
wi_pr_ac <- split(x = wi_pr_ac, f = wi_pr_ac$anno)
wi_pr_ac <- wi_pr_ac[lapply(wi_pr_ac, nrow) > 10]
wi_proximal_active <- plyr::ldply(wi_pr_ac, data.frame)[, -1]
wi_k_ac <- gsub(pattern = "proximal.active.", replacement = "", x = wi_proximal_active$anno)
wi_k_ac <- change_names(wi_k_ac)

wi_k_ac <- factor(wi_k_ac, unique(wi_k_ac)[c(7, 9, 8, 3, 6, 4, 5, 1, 2)])

draw_EH(x = wi_proximal_active$Name, k = wi_k_ac, name = "proximal_active", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_proximal_active$Name, k = wi_k_ac, name = "proximal_active", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_proximal_active$Name, k = wi_k_ac, name = "proximal_active", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_proximal_active$Name, k = wi_k_ac, name = "proximal_active", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```

### Proximal inactive regions
```{r enriched-heatmaps-data-overlap-22 }
wi_pr_inac <- wi[grep(pattern = "proximal.inactive", x = wi$anno), ]
wi_pr_inac <- split(x = wi_pr_inac, f = wi_pr_inac$anno)
wi_pr_inac <- wi_pr_inac[lapply(wi_pr_inac, nrow) > 10]
wi_proximal_inactive <- plyr::ldply(wi_pr_inac, data.frame)[, -1]
wi_k_inac <- gsub(pattern = "proximal.inactive.", replacement = "", x = wi_proximal_inactive$anno)
wi_k_inac <- change_names(n = wi_k_inac)

wi_k_inac <- factor(wi_k_inac, unique(wi_k_inac)[c(1:3, 6, 4, 5)])

draw_EH(x = wi_proximal_inactive$Name, k = wi_k_inac, name = "proximal_inactive", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "none")
draw_EH(x = wi_proximal_inactive$Name, k = wi_k_inac, name = "proximal_inactive", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "none")
draw_EH(x = wi_proximal_inactive$Name, k = wi_k_inac, name = "proximal_inactive", ht = 14.5, type = "wi", log = FALSE, outdir = "with_ChIP", rm = "meth")
draw_EH(x = wi_proximal_inactive$Name, k = wi_k_inac, name = "proximal_inactive", ht = 14.5, type = "wi", log = TRUE, outdir = "with_ChIP_log2", rm = "meth")
```


# References
```{r enriched-heatmaps-data-overlap-23 }
report::cite_packages(sessionInfo())
```


# SessionInfo
```{r enriched-heatmaps-data-overlap-24 }
devtools::session_info()
```
