---
title: "ATAC-Seq: Peak annotation and figures"
author: "Deepak Tanwar"
date: "<b>Created on:</b> 2019-11-18 <br> <b>Updated on:</b> `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: tango
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    keep_md: no
    number_sections: no
    fig_width: 8
    fig_height: 8
    fig_caption: true
    df_print: paged
    code_folding: hide
  fontsize: 12pt
  geometry: margin=1in
  documentclass: article
# bibliography: references.bib
link-citations: yes
editor_options: 
  chunk_output_type: console
---

# Library
```{r 20191118-peaks-annotation-other-figs-1, message=FALSE, warning=FALSE}
library(csaw)
library(edgeR)
library(ChIPseeker)
library(ggplotify)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
library(EnrichedHeatmap)
```

# TxDb and Promoter
```{r 20191118-peaks-annotation-other-figs-2 }
txdb <- TxDb.Mmusculus.UCSC.mm10.knownGene
promoter <- getPromoters(TxDb = txdb, upstream = 1000, downstream = 1000)
regions <- getPromoters(TxDb = txdb, upstream = 5000, downstream = 5000)
```

# All peaks
## Reading the peak file
```{r 20191118-peaks-annotation-other-figs-3 }
allAtac <- GRanges(read.table("input/PND_Adult_NF_peaks.narrowPeak",
  header = F,
  col.names = c(
    "Chr", "Start", "End", "Name", "Score", "Strand",
    "FoldChange", "negLog10pvalue", "negLog10qvalue",
    "Summit"
  ), stringsAsFactors = F
))

PNDAtac <- GRanges(read.table("input/PND15_NF_peaks.narrowPeak",
  header = F,
  col.names = c(
    "Chr", "Start", "End", "Name", "Score", "Strand",
    "FoldChange", "negLog10pvalue", "negLog10qvalue",
    "Summit"
  ), stringsAsFactors = F
))

AdultAtac <- GRanges(read.table("input/Adult_NF_peaks.narrowPeak",
  header = F,
  col.names = c(
    "Chr", "Start", "End", "Name", "Score", "Strand",
    "FoldChange", "negLog10pvalue", "negLog10qvalue",
    "Summit"
  ), stringsAsFactors = F
))
```

## Analysis
```{r 20191118-peaks-annotation-other-figs-4, message=FALSE, warning=FALSE}
tagMatrix.all <- getTagMatrix(allAtac, windows = regions)

tagPlot <- as.ggplot(~ tagHeatmap(
    tagMatrix = tagMatrix.all, xlim = c(-5000, 5000),
    color = "red", title = "All peaks"
))

peakAnno.all <- annotatePeak(peak = allAtac, 
                             TxDb = txdb, 
                             tssRegion = c(-1000, 1000), 
                             annoDb = "org.Mm.eg.db")

pie <- as.ggplot(~plotAnnoPie(peakAnno.all))
barPlot <- plotAnnoBar(peakAnno.all)
tssPlot <- plotDistToTSS(peakAnno.all)
upset <- upsetplot(peakAnno.all, vennpie = TRUE)
avgProfile <- plotAvgProf(tagMatrix.all, xlim = c(-5000, 5000), conf = 0.95, resample = 100)

allPeaks <- list(tagMatrix = tagMatrix.all,
                 peakAnno = peakAnno.all,
                 tagsPlot = tagPlot,
                 barPlot = barPlot,
                 tssPlot = tssPlot,
                 upsetPlot = upset,
                 avgProfile = avgProfile,
                 pieChart = pie)
```

## TagMatrix Plot
```{r 20191118-peaks-annotation-other-figs-5, fig.align='center', fig.height=11, fig.width=6}
allPeaks$tagsPlot
```

## Pie chart
```{r 20191118-peaks-annotation-other-figs-6, fig.width=8, fig.height=8, fig.align='center'}
allPeaks$pieChart
```

## Barplot
```{r 20191118-peaks-annotation-other-figs-7, fig.width=8, fig.height=3, fig.align='center'}
allPeaks$barPlot
```

## Distance to TSS plot
```{r 20191118-peaks-annotation-other-figs-8, fig.width=8, fig.height=3, fig.align='center'}
allPeaks$tssPlot
```

## UpSet plot
```{r 20191118-peaks-annotation-other-figs-9, warning=FALSE, message=FALSE, fig.align='center'}
allPeaks$upset
```

## Average Profile of ATAC peaks
```{r 20191118-peaks-annotation-other-figs-10, fig.align='center', fig.width=8, fig.height=7, warning=FALSE, message=FALSE}
allPeaks$avgProfile
```

# Differential accessible regions (peaks)

## Reading the data
```{r 20191118-peaks-annotation-other-figs-11 }
load("./input/atac_diff_chromatin_accessibility.RData")
da <- data.frame(rowData(data), stringsAsFactors = F, check.names = F)
da <- da[(da$`diffAccessibility-logFC` >= 1 | da$`diffAccessibility-logFC` <= -1) &
           da$`diffAccessibility-qvalue` <= 0.05,]

colnames(da)[1:5] <- gsub(pattern = "Peak-", replacement = "", x = colnames(da)[1:5])

da.GR <- GRanges(da)
```

## Differential Accessible regions (all)

### Analysis
```{r 20191118-peaks-annotation-other-figs-12, message=FALSE, warning=FALSE}
tagMatrix.da <- getTagMatrix(da.GR, windows = regions)

tagPlot <- as.ggplot(~ tagHeatmap(
    tagMatrix = tagMatrix.da, xlim = c(-5000, 5000),
    color = "red", title = "Differential accessible peaks"
))

peakAnno.da <- annotatePeak(peak = da.GR, 
                             TxDb = txdb, 
                             tssRegion = c(-1000, 1000), 
                             annoDb = "org.Mm.eg.db")

pie <- as.ggplot(~plotAnnoPie(peakAnno.da))
barPlot <- plotAnnoBar(peakAnno.da)
tssPlot <- plotDistToTSS(peakAnno.da)
upset <- upsetplot(peakAnno.da, vennpie = TRUE)
avgProfile <- plotAvgProf(tagMatrix.da, xlim = c(-5000, 5000), conf = 0.95, resample = 100)

daPeaks <- list(tagMatrix = tagMatrix.da,
                 peakAnno = peakAnno.da,
                 tagsPlot = tagPlot,
                 barPlot = barPlot,
                 tssPlot = tssPlot,
                 upsetPlot = upset,
                 avgProfile = avgProfile,
                 pieChart = pie)
```

### TagMatrix Plot
```{r 20191118-peaks-annotation-other-figs-13, fig.align='center', fig.height=11, fig.width=6}
daPeaks$tagsPlot
```

### Pie chart
```{r 20191118-peaks-annotation-other-figs-14, fig.width=8, fig.height=8, fig.align='center'}
daPeaks$pieChart
```

### Barplot
```{r 20191118-peaks-annotation-other-figs-15, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$barPlot
```

### Distance to TSS plot
```{r 20191118-peaks-annotation-other-figs-16, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$tssPlot
```

### UpSet plot
```{r 20191118-peaks-annotation-other-figs-17, warning=FALSE, message=FALSE, fig.align='center'}
daPeaks$upset
```

### Average Profile of ATAC peaks
```{r 20191118-peaks-annotation-other-figs-18, fig.align='center', fig.width=8, fig.height=7, warning=FALSE, message=FALSE}
daPeaks$avgProfile
```


## Differential Accessible regions (Positive in Adults)

### Analysis
```{r 20191118-peaks-annotation-other-figs-19, message=FALSE, warning=FALSE}
tagMatrix.da <- getTagMatrix(da.GR[da.GR$`diffAccessibility-logFC` > 0,], 
                             windows = regions)

tagPlot <- as.ggplot(~ tagHeatmap(
    tagMatrix = tagMatrix.da, xlim = c(-5000, 5000),
    color = "red", title = "Differential accessible peaks (Positive in Adults)"
))

peakAnno.da <- annotatePeak(peak = da.GR, 
                             TxDb = txdb, 
                             tssRegion = c(-1000, 1000), 
                             annoDb = "org.Mm.eg.db")

pie <- as.ggplot(~plotAnnoPie(peakAnno.da))
barPlot <- plotAnnoBar(peakAnno.da)
tssPlot <- plotDistToTSS(peakAnno.da)
upset <- upsetplot(peakAnno.da, vennpie = TRUE)
avgProfile <- plotAvgProf(tagMatrix.da, xlim = c(-5000, 5000), conf = 0.95, resample = 100)

daPeaks <- list(tagMatrix = tagMatrix.da,
                 peakAnno = peakAnno.da,
                 tagsPlot = tagPlot,
                 barPlot = barPlot,
                 tssPlot = tssPlot,
                 upsetPlot = upset,
                 avgProfile = avgProfile,
                 pieChart = pie)
```

### TagMatrix Plot
```{r 20191118-peaks-annotation-other-figs-20, fig.align='center', fig.height=11, fig.width=6}
daPeaks$tagsPlot
```

### Pie chart
```{r 20191118-peaks-annotation-other-figs-21, fig.width=8, fig.height=8, fig.align='center'}
daPeaks$pieChart
```

### Barplot
```{r 20191118-peaks-annotation-other-figs-22, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$barPlot
```

### Distance to TSS plot
```{r 20191118-peaks-annotation-other-figs-23, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$tssPlot
```

### UpSet plot
```{r 20191118-peaks-annotation-other-figs-24, warning=FALSE, message=FALSE, fig.align='center'}
daPeaks$upset
```

### Average Profile of ATAC peaks
```{r 20191118-peaks-annotation-other-figs-25, fig.align='center', fig.width=8, fig.height=7, warning=FALSE, message=FALSE}
daPeaks$avgProfile
```


## Differential Accessible regions (Negative in Adults)

### Analysis
```{r 20191118-peaks-annotation-other-figs-26, message=FALSE, warning=FALSE}
tagMatrix.da <- getTagMatrix(da.GR[da.GR$`diffAccessibility-logFC` > 0,], 
                             windows = regions)

tagPlot <- as.ggplot(~ tagHeatmap(
    tagMatrix = tagMatrix.da, xlim = c(-5000, 5000),
    color = "red", title = "Differential accessible peaks (Negative in Adults)"
))

peakAnno.da <- annotatePeak(peak = da.GR, 
                             TxDb = txdb, 
                             tssRegion = c(-1000, 1000), 
                             annoDb = "org.Mm.eg.db")

pie <- as.ggplot(~plotAnnoPie(peakAnno.da))
barPlot <- plotAnnoBar(peakAnno.da)
tssPlot <- plotDistToTSS(peakAnno.da)
upset <- upsetplot(peakAnno.da, vennpie = TRUE)
avgProfile <- plotAvgProf(tagMatrix.da, xlim = c(-5000, 5000), conf = 0.95, resample = 100)

daPeaks <- list(tagMatrix = tagMatrix.da,
                 peakAnno = peakAnno.da,
                 tagsPlot = tagPlot,
                 barPlot = barPlot,
                 tssPlot = tssPlot,
                 upsetPlot = upset,
                 avgProfile = avgProfile,
                 pieChart = pie)
```

### TagMatrix Plot
```{r 20191118-peaks-annotation-other-figs-27, fig.align='center', fig.height=11, fig.width=6}
daPeaks$tagsPlot
```

### Pie chart
```{r 20191118-peaks-annotation-other-figs-28, fig.width=8, fig.height=8, fig.align='center'}
daPeaks$pieChart
```

### Barplot
```{r 20191118-peaks-annotation-other-figs-29, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$barPlot
```

### Distance to TSS plot
```{r 20191118-peaks-annotation-other-figs-30, fig.width=8, fig.height=3, fig.align='center'}
daPeaks$tssPlot
```

### UpSet plot
```{r 20191118-peaks-annotation-other-figs-31, warning=FALSE, message=FALSE, fig.align='center'}
daPeaks$upset
```

### Average Profile of ATAC peaks
```{r 20191118-peaks-annotation-other-figs-32, fig.align='center', fig.width=8, fig.height=7, warning=FALSE, message=FALSE}
daPeaks$avgProfile
```


# SessionInfo
```{r 20191118-peaks-annotation-other-figs-33 }
devtools::session_info()
```
