---
title: "RNA-Seq: Pathway analysis using cameraPR and fGSEA"
author: "Deepak Tanwar"
date: "<b>Created on:</b> 2019-09-02 <br> <b>Updated on:</b> `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: tango
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    keep_md: no
    number_sections: no
    fig_width: 8
    fig_height: 8
    fig_caption: true
    df_print: paged
    code_folding: hide
  fontsize: 12pt
  geometry: margin=1in
  documentclass: article
# bibliography: references.bib
link-citations: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries required

```{r pathways-analysis-1, message=FALSE, warning=FALSE}
library(fgsea)
library(limma)
library(parallel)
```

# Data setup
```{r pathways-analysis-2 }
load("input/limma_SC_Controls_lab_lit.RData")
load("input/gencode.vM18.anno.RData")
```


# Genes with length information
```{r pathways-analysis-3 }
genes <- anno[, c(2, 3, 6, 7, 10)]
genes <- genes[!duplicated(genes), ]

g1 <- genes[!genes$entrez_id %in% genes$entrez_id[duplicated(genes$entrez_id)], ]
g2 <- genes[genes$entrez_id %in% genes$entrez_id[duplicated(genes$entrez_id)], ]

g2 <- g2[!duplicated(g2$ensembl), ]
g2 <- g2[!duplicated(g2$symbol), ]
g2$entrez_id[duplicated(g2$entrez_id)] <- g2$symbol[duplicated(g2$entrez_id)]

genes.f <- rbind(g1, g2)
rownames(genes.f) <- genes.f$entrez_id
```


# GeneSets
```{r pathways-analysis-4 }
if (!file.exists("output/pathways.rds")) {
  MSigdb <- getMSigGeneSetDb(
    collection = c("c1", "c2", "c3", "c4", "c5", "c6", "c7", "h"),
    species = "mouse", with.kegg = T, species.specific = T
  )

  go <- GeneSetDb(getGenesets(org = "mmu", db = "go"), collectionName = "Gene Ontology")
  go.bp <- GeneSetDb(getGenesets(org = "mmu", go.onto = "BP"), collectionName = "GO_BP")
  go.mf <- GeneSetDb(getGenesets(org = "mmu", go.onto = "MF"), collectionName = "GO_MF")
  go.cc <- GeneSetDb(getGenesets(org = "mmu", go.onto = "CC"), collectionName = "GO_CC")

  kegg <- GeneSetDb(getGenesets(org = "mmu", db = "kegg"), collectionName = "KEGG")

  reactome <- getReactomeGeneSetDb(species = "mouse", rm.species.prefix = T)

  gdb <- Reduce(append, list(go, go.bp, go.cc, go.mf, kegg, reactome, MSigdb))
  gdb@table$organism <- "Mus musculus"

  pathwaysDB <- data.frame(gdb@db, stringsAsFactors = F)
  pathwaysDB$symbol_anno <- genes.f[pathwaysDB$featureId, "symbol"]
  pathwaysDB$symbol_anno[is.na(pathwaysDB$symbol_anno)] <- pathwaysDB$featureId[is.na(pathwaysDB$symbol_anno)]

  pathways <- split(x = pathwaysDB[, c(1, 2, 5)], f = pathwaysDB$collection)
  pathways <- lapply(pathways, function(x) split(x[, 3], x$name))

  saveRDS(object = pathways, file = "output/pathways.rds")
} else {
  pathways <- readRDS("output/pathways.rds")
}

pathways <- pathways[5:7]
```


# fgsea
```{r pathways-analysis-5 }
fgsea.pnd8.pnd15 <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd8 vs pnd15`
  genes <- dea$t
  names(genes) <- dea$Genes
  res <- fgsea(pathways = x, stats = genes, nperm = 100000, minSize = 10, maxSize = 1000, nproc = detectCores() - 1)
  # res <- data.frame(res[res$padj <= 0.05, ])
  rownames(res) <- res$pathway
  return(res)
})

fgsea.pnd15.adult <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd14 vs pnw8`
  genes <- dea$logFC
  names(genes) <- dea$Genes
  res <- fgsea(pathways = x, stats = genes, nperm = 100000, minSize = 10, maxSize = 1000, nproc = detectCores() - 1)
  # res <- data.frame(res[res$padj <= 0.05, ])
  rownames(res) <- res$pathway
  return(res)
})
```

# fgsea Multi-level
```{r pathways-analysis-6 }
fgsea.pnd8.pnd15_m <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd8 vs pnd15`
  genes <- dea$t
  names(genes) <- dea$Genes
  res <- fgseaMultilevel(pathways = x, stats = genes, minSize = 10, maxSize = 1000, eps = 1e-100, nproc = detectCores() - 1)
  # res <- data.frame(res[res$padj <= 0.05, ])
  rownames(res) <- res$pathway
  return(res)
})

fgsea.pnd15.adult_m <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd14 vs pnw8`
  genes <- dea$logFC
  names(genes) <- dea$Genes
  res <- fgseaMultilevel(pathways = x, stats = genes, minSize = 10, maxSize = 1000, eps = 1e-100, nproc = detectCores() - 1)
  # res <- data.frame(res[res$padj <= 0.05, ])
  rownames(res) <- res$pathway
  return(res)
})
```


# cameraPR

## Function
```{r pathways-analysis-7 }
cameraPR <- function(statistic, index, use.ranks = FALSE, inter.gene.cor = 0.01, sort = TRUE, ...) {
  dots <- names(list(...))
  if (length(dots)) warning("Extra arguments disregarded: ", sQuote(dots))

  #     Check statistic
  if (is.list(statistic)) stop("statistic should be a numeric vector")
  storage.mode(statistic) <- "numeric"
  if (anyNA(statistic)) stop("NA values for statistic not allowed")
  G <- length(statistic)
  ID <- names(statistic)
  if (G < 3) stop("Too few genes in dataset: need at least 3")

  #     Check index
  if (!is.list(index)) index <- list(set1 = index)
  index <- index[which(!sapply(index, function(x) length(x) < 10))] # ADDED BY ASA 190527 to mimic fgsea minSize = 10
  nsets <- length(index)

  #     Check inter.gene.cor
  if (anyNA(inter.gene.cor)) stop("NA inter.gene.cor not allowed")
  if (any(abs(inter.gene.cor) >= 1)) stop("inter.gene.cor too large or small")
  if (length(inter.gene.cor) > 1L) {
    if (length(inter.gene.cor) != nsets) stop("Length of inter.gene.cor doesn't match number of sets")
    fixed.cor <- FALSE
  } else {
    fixed.cor <- TRUE
    inter.gene.cor <- rep_len(inter.gene.cor, nsets)
  }

  #     Set df
  if (use.ranks) {
    df.camera <- Inf
  } else {
    df.camera <- G - 2L
  }

  #     Global statistics
  meanStat <- mean(statistic)
  varStat <- var(statistic)
  tst <- c() ### ADDED BY ASA on 190514
  NGenes <- Down <- Up <- rep_len(0, nsets)
  for (i in 1:nsets) {
    iset <- index[[i]]
    if (is.character(iset)) iset <- which(ID %in% iset)
    StatInSet <- statistic[iset]
    m <- length(StatInSet)
    NGenes[i] <- m
    if (use.ranks) {
      p.value <- rankSumTestWithCorrelation(iset,
        statistics = statistic,
        correlation = inter.gene.cor[i], df = df.camera
      )
      Down[i] <- p.value[1]
      Up[i] <- p.value[2]
    } else {
      vif <- 1 + (m - 1) * inter.gene.cor[i]
      m2 <- G - m
      meanStatInSet <- mean(StatInSet)
      delta <- G / m2 * (meanStatInSet - meanStat)
      varStatPooled <- ((G - 1L) * varStat - delta^2 * m * m2 / G) / (G - 2L)
      two.sample.t <- delta / sqrt(varStatPooled * (vif / m + 1 / m2))
      tst[i] <- two.sample.t ### ADDED BY ASA on 190514
      Down[i] <- pt(two.sample.t, df = df.camera)
      Up[i] <- pt(two.sample.t, df = df.camera, lower.tail = FALSE)
    }
  }
  TwoSided <- 2 * pmin(Down, Up)
  #     Assemble into data.frame
  D <- (Down < Up)
  Direction <- rep_len("Up", nsets)
  Direction[D] <- "Down"
  if (fixed.cor) {
    tab <- data.frame(NGenes = NGenes, Direction = Direction, PValue = TwoSided, TwoSampT = tst, stringsAsFactors = FALSE)
  } ### ADDED TwoSampT on 190514
  else {
    tab <- data.frame(NGenes = NGenes, Correlation = inter.gene.cor, Direction = Direction, PValue = TwoSided, stringsAsFactors = FALSE)
  }
  rownames(tab) <- names(index)

  #     Add FDR
  if (nsets > 1L) tab$FDR <- p.adjust(tab$PValue, method = "BH")

  #     Sort by p-value
  if (sort && nsets > 1L) {
    o <- order(tab$PValue)
    tab <- tab[o, ]
  }

  tab
}
```

## cameraPR
```{r pathways-analysis-8 }
cameraPR.pnd8.pnd15 <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd8 vs pnd15`
  genes <- dea$t
  names(genes) <- dea$Genes
  res <- cameraPR(genes, x)
  # res <- res[res$FDR <= 0.05, ]
  res <- data.frame(pathway = rownames(res), res, stringsAsFactors = F)
  return(res)
})

cameraPR.pnd15.adult <- lapply(pathways, function(x) {
  dea <- dea.limma$`pnd14 vs pnw8`
  genes <- dea$logFC
  names(genes) <- dea$Genes
  res <- cameraPR(genes, x)
  # res <- res[res$FDR <= 0.05, ]
  res <- data.frame(pathway = rownames(res), res, stringsAsFactors = F)
  return(res)
})
```

# Tables

## PND8 vs PND15

### fGSEA

```{r pathways-analysis-11 }
fgsea.pnd8.pnd15 <- lapply(fgsea.pnd8.pnd15, function(x) {
  x$leadingEdgeGenes <- NA
  for (i in 1:nrow(x)) {
    x$leadingEdgeGenes[i] <- paste(x$leadingEdge[[i]], collapse = ",")
  }
  x <- data.frame(x)
  x <- x[, !colnames(x) %in% "leadingEdge"]
  return(x)
})

path <- fgsea.pnd8.pnd15

for (i in 1:length(path)) {
  n <- paste0("./output/fGSEA_PND8vsPND15_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = data.frame(path[[i]]), path = n, col_names = T, format_headers = T)
}
```


### fGSEA multilevel
```{r pathways-analysis-12 }
fgsea.pnd8.pnd15_m <- lapply(fgsea.pnd8.pnd15_m, function(x) {
  x$leadingEdgeGenes <- NA
  for (i in 1:nrow(x)) {
    x$leadingEdgeGenes[i] <- paste(x$leadingEdge[[i]], collapse = ",")
  }
  x <- data.frame(x)
  x <- x[, !colnames(x) %in% "leadingEdge"]
  return(x)
})

path <- fgsea.pnd8.pnd15_m

for (i in 1:length(path)) {
  n <- paste0("./output/fGSEA_multilevel_PND8vsPND15_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = data.frame(path[[i]]), path = n, col_names = T, format_headers = T)
}
```



### CameraPR

```{r pathways-analysis-13 }
path <- cameraPR.pnd8.pnd15

for (i in 1:length(path)) {
  n <- paste0("./output/cameraPR_PND8vsPND15_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = path[[i]], path = n, col_names = T, format_headers = T)
}
```


## PND14 vs PNW8

### fGSEA

```{r pathways-analysis-14 }
fgsea.pnd15.adult <- lapply(fgsea.pnd15.adult, function(x) {
  x$leadingEdgeGenes <- NA
  for (i in 1:nrow(x)) {
    x$leadingEdgeGenes[i] <- paste(x$leadingEdge[[i]], collapse = ",")
  }
  x <- data.frame(x)
  x <- x[, !colnames(x) %in% "leadingEdge"]
  return(x)
})

path <- fgsea.pnd15.adult

for (i in 1:length(path)) {
  n <- paste0("./output/fGSEA_PND14vsPNW8_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = path[[i]], path = n, col_names = T, format_headers = T)
}
```


### fGSEA multilevel

```{r pathways-analysis-15 }
fgsea.pnd15.adult_m <- lapply(fgsea.pnd15.adult_m, function(x) {
  x$leadingEdgeGenes <- NA
  for (i in 1:nrow(x)) {
    x$leadingEdgeGenes[i] <- paste(x$leadingEdge[[i]], collapse = ",")
  }
  x <- data.frame(x)
  x <- x[, !colnames(x) %in% "leadingEdge"]
  return(x)
})

path <- fgsea.pnd15.adult_m

for (i in 1:length(path)) {
  n <- paste0("./output/fGSEA_multilevel_PND14vsPNW8_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = path[[i]], path = n, col_names = T, format_headers = T)
}
```



### CameraPR

```{r pathways-analysis-16 }
path <- cameraPR.pnd15.adult

for (i in 1:length(path)) {
  n <- paste0("./output/cameraPR_PND14vsPNW8_", names(path)[i], ".xlsx")
  n <- gsub(pattern = " ", replacement = "-", x = n)
  writexl::write_xlsx(x = path[[i]], path = n, col_names = T, format_headers = T)
}
```


# Common pathways


## PND8 vs PND15

### fGSEA + cameraPR
```{r pathways-analysis-17 }
for (i in 1:length(fgsea.pnd8.pnd15)) {
  n <- names(fgsea.pnd8.pnd15)[i]
  tab1 <- fgsea.pnd8.pnd15[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- cameraPR.pnd8.pnd15[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("cameraPR", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_cameraPR_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```

### fGSEA multilevel + cameraPR
```{r pathways-analysis-18 }
for (i in 1:length(fgsea.pnd8.pnd15_m)) {
  n <- names(fgsea.pnd8.pnd15_m)[i]
  tab1 <- fgsea.pnd8.pnd15_m[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- cameraPR.pnd8.pnd15[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("cameraPR", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEAmultiLevel_cameraPR_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```


### fGSEA + fGSEA multilevel
```{r pathways-analysis-19 }
for (i in 1:length(fgsea.pnd8.pnd15)) {
  n <- names(fgsea.pnd8.pnd15)[i]
  tab1 <- fgsea.pnd8.pnd15[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- fgsea.pnd8.pnd15_m[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("fGSEAmultiLevel", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_fGSEAmultiLevel_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```

### fGSEA + fGSEA multilevel + cameraPR
```{r pathways-analysis-20 }
for (i in 1:length(fgsea.pnd8.pnd15)) {
  n <- names(fgsea.pnd8.pnd15)[i]
  tab1 <- fgsea.pnd8.pnd15[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- fgsea.pnd8.pnd15_m[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("fGSEAmultiLevel", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab3 <- cameraPR.pnd8.pnd15[[n]]
  colnames(tab3)[2:ncol(tab3)] <- paste("cameraPR", colnames(tab3)[2:ncol(tab3)], sep = "-")

  tab <- merge(tab1, tab2, by = "pathway")
  tab <- merge(tab, tab3, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_fGSEAmultiLevel_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```



## PND15 vs Adult

### fGSEA + cameraPR
```{r pathways-analysis-21 }
for (i in 1:length(fgsea.pnd15.adult)) {
  n <- names(fgsea.pnd15.adult)[i]
  tab1 <- fgsea.pnd15.adult[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- cameraPR.pnd15.adult[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("cameraPR", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_cameraPR_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```

### fGSEA multilevel + cameraPR
```{r pathways-analysis-22 }
for (i in 1:length(fgsea.pnd15.adult_m)) {
  n <- names(fgsea.pnd15.adult_m)[i]
  tab1 <- fgsea.pnd15.adult_m[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- cameraPR.pnd15.adult[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("cameraPR", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEAmultiLevel_cameraPR_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```


### fGSEA + fGSEA multilevel
```{r pathways-analysis-23 }
for (i in 1:length(fgsea.pnd15.adult)) {
  n <- names(fgsea.pnd15.adult)[i]
  tab1 <- fgsea.pnd15.adult[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- fgsea.pnd15.adult_m[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("fGSEAmultiLevel", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab <- merge(tab1, tab2, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_fGSEAmultiLevel_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```

### fGSEA + fGSEA multilevel + cameraPR
```{r pathways-analysis-24 }
for (i in 1:length(fgsea.pnd15.adult)) {
  n <- names(fgsea.pnd15.adult)[i]
  tab1 <- fgsea.pnd15.adult[[n]]
  colnames(tab1)[2:ncol(tab1)] <- paste("fGSEA", colnames(tab1)[2:ncol(tab1)], sep = "-")
  tab2 <- fgsea.pnd15.adult_m[[n]]
  colnames(tab2)[2:ncol(tab2)] <- paste("fGSEAmultiLevel", colnames(tab2)[2:ncol(tab2)], sep = "-")
  tab3 <- cameraPR.pnd15.adult[[n]]
  colnames(tab3)[2:ncol(tab3)] <- paste("cameraPR", colnames(tab3)[2:ncol(tab3)], sep = "-")

  tab <- merge(tab1, tab2, by = "pathway")
  tab <- merge(tab, tab3, by = "pathway")

  writexl::write_xlsx(
    x = tab, path = paste0("./output/fGSEA_fGSEAmultiLevel_PND8vsPND15_", n, ".xlsx"),
    col_names = T, format_headers = T
  )
}
```

# References
```{r pathways-analysis-25 }
report::cite_packages(sessionInfo())
```


# SessionInfo
```{r pathways-analysis-26 }
devtools::session_info()
```
