---
title: "RNA-Seq: differential expression analysis of controls"
author: "Deepak Tanwar"
date: "<b>Created on:</b> 2019-05-13 <br> <b>Updated on:</b> `r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: tango
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    keep_md: no
    number_sections: no
    fig_width: 8
    fig_height: 8
    fig_caption: true
    df_print: paged
    code_folding: show
  fontsize: 12pt
  geometry: margin=1in
  documentclass: article
# bibliography: references.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries required

```{r, message=FALSE, warning=FALSE}
library(rmarkdown)
library(limma)
library(edgeR)
library(knitr)
library(plgINS)
library(fdrtool)
library(plyr)
library(miceadds)
library(forcats)
```

# Function to change names of PND data
```{r}
change_names <- function(names) {
  names <- gsub(pattern = "PND7", replacement = "PND8", x = names)
  names <- gsub(pattern = "PND14", replacement = "PND15", x = names)
  names <- gsub(pattern = "_C", replacement = "_CTRL_", x = names)
  names <- as.character(sapply(names, function(x) {
    r <- NULL
    if (substr(x = x, 1, 4) == "mRNA") {
      r <- paste(substr(x, 6, 20), substr(x, 1, 4), sep = "_")
    } else {
      r <- x
    }
    return(r)
  }))
  return(names)
}
```

# Salmon data objects

## Load the data and filter it
```{r}
pnd <- load.Rdata2("input/salmon_msus27_SSC.tds.RData")
pnd <- pnd[pnd@phenoData$group == "C"]
colnames(pnd@gene.counts) <- change_names(names = colnames(pnd@gene.counts))

adult <- load.Rdata2("input/salmon_msus30_SC_adult.tds.RData")
adult <- adult[adult@phenoData$Group == "CTRL"]

counts <- data.matrix(data.frame(pnd@gene.counts, adult@gene.counts[rownames(pnd@gene.counts), ]))

pData.pnd <- pnd@phenoData[, c(1, 2, 4)]
colnames(pData.pnd) <- c("Age", "Batch", "Cage.ID")
pData.pnd$Age[pData.pnd$Age == 7] <- 8
pData.pnd$Age[pData.pnd$Age == 14] <- 15
pData.pnd$Group <- paste0("PND", pData.pnd$Age)
rownames(pData.pnd) <- change_names(names = rownames(pData.pnd))

pData.adult <- adult@phenoData[3]
pData.adult$Age <- 150
pData.adult$Batch <- "D"
pData.adult$Group <- "Adult"

pData <- rbind.data.frame(pData.pnd, pData.adult, stringsAsFactors = F)
pData$Age <- as.factor(pData$Age)
```

## New salmon object for SC Controls
```{r}
sc <- pnd
sc@phenoData <- pData
sc@norm.factors <- calcNormFactors(counts)

colnames(sc@tx.counts) <- change_names(names = colnames(sc@tx.counts))
sc@tx.counts <- data.matrix(data.frame(sc@tx.counts, adult@tx.counts[rownames(sc@tx.counts), ]))
sc@tx.counts <- sc@tx.counts[, rownames(pData)]

colnames(sc@tx.tpm) <- change_names(names = colnames(sc@tx.tpm))
sc@tx.tpm <- data.matrix(data.frame(sc@tx.tpm, adult@tx.tpm[rownames(sc@tx.tpm), ]))
sc@tx.tpm <- sc@tx.tpm[, rownames(pData)]

sc@gene.counts <- counts

colnames(sc@gene.tpm) <- change_names(names = colnames(sc@gene.tpm))
sc@gene.tpm <- data.matrix(data.frame(sc@gene.tpm, adult@gene.tpm[rownames(sc@gene.tpm), ]))
sc@gene.tpm <- sc@gene.tpm[, rownames(pData)]

colnames(sc@type.readProp) <- change_names(names = colnames(sc@type.readProp))
sc@type.readProp <- data.matrix(data.frame(sc@type.readProp, adult@type.readProp[rownames(sc@type.readProp), ]))
sc@type.readProp <- sc@type.readProp[, rownames(pData)]

colnames(sc@type.tpm) <- change_names(names = colnames(sc@type.tpm))
sc@type.tpm <- data.matrix(data.frame(sc@type.tpm, adult@type.tpm[rownames(sc@type.tpm), ]))
sc@type.tpm <- sc@type.tpm[, rownames(pData)]

rownames(sc@type.tmm)[15] <- "repeat_family"
colnames(sc@type.tmm) <- change_names(names = colnames(sc@type.tmm))
sc@type.tmm <- data.matrix(data.frame(sc@type.tmm, adult@type.tmm[rownames(sc@type.tmm), ]))
sc@type.tmm <- sc@type.tmm[, rownames(pData)]

sc@libStats <- rbind(sc@libStats, adult@libStats[, -5])
sc@cmd <- rbind(sc@cmd, adult@cmd[, -6])

salmon <- sc
save(object = salmon, file = "./output/salmon_SC_controls.tds.RData")
```


# Filtering gene counts

## Function
```{r}
counts_filter <- function(counts, nReads = 15, nSamplesPercent = 0.4) {
  counts2 <- counts[apply(counts, 1, FUN = function(x) {
    sum(x >= 15) >= floor(length(x) * 0.4)
  }), ]
  return(counts2)
}
```

## Counts from all data
```{r}
counts.filt <- counts_filter(counts = counts)
```


# Differeitial analysis using `limma`

## Function for differential analysis using limma
```{r}
voomDupCor <- function(counts, mm, individuals, tcoefs = NULL) {
  ## model matrix = model.matrix(~batch+group, data = desc)
  ## individuals = sample/cage column from desc
  if (is.null(tcoefs)) tcoefs <- colnames(mm)[ncol(mm)]
  dds <- calcNormFactors(DGEList(counts))
  v <- voom(dds, mm)
  dc <- duplicateCorrelation(v, mm, block = individuals)
  v <- voom(dds, mm, block = individuals, correlation = dc$consensus.correlation)
  fit <- eBayes(lmFit(v, mm, block = individuals, correlation = dc$consensus.correlation))
  tags <- as.data.frame(topTable(fit, tcoefs, 60000))
  tags <- data.frame(Genes = rownames(tags), tags, stringsAsFactors = F)
  return(list(DEA = tags, voom.norm.counts = v$E, EList = v))
}

voomDupCor.s <- function(counts, mm, tcoefs = NULL) {
  if (is.null(tcoefs)) tcoefs <- colnames(mm)[ncol(mm)]
  dds <- calcNormFactors(DGEList(counts))
  v <- voom(dds, mm)
  fit <- eBayes(lmFit(v, mm))
  tags <- as.data.frame(topTable(fit, tcoefs, 60000))
  tags <- data.frame(Genes = rownames(tags), tags, stringsAsFactors = F)
  return(list(DEA = tags, voom.norm.counts = v$E, EList = v))
}
```

## Function for running differential analysis
```{r}
run_DEA <- function(mat, pData, groupCol = "Group", group1, group2, batch = NULL, cage = NULL, revGroup = F) {
  pheno <- pData[pData[, groupCol] %in% c(group1, group2), ]

  if (revGroup) pheno[, groupCol] <- fct_rev(pheno[, groupCol])

  tab <- mat[, rownames(pheno)]

  dea <- NULL

  if (!is.null(cage)) {
    if (!is.null(batch)) {
      mm <- model.matrix(~ pheno[, batch] + pheno[, groupCol])
      dea <- voomDupCor(counts = tab, mm = mm, individuals = pheno[, cage])
    } else {
      mm <- model.matrix(~ pheno[, groupCol])
      dea <- voomDupCor(counts = tab, mm = mm, individuals = pheno[, cage])
    }
  } else {
    if (!is.null(batch)) {
      mm <- model.matrix(~ pheno[, batch] + pheno[, groupCol])
      dea <- voomDupCor.s(counts = tab, mm = mm)
    } else {
      mm <- model.matrix(~ pheno[, groupCol])
      dea <- voomDupCor.s(counts = tab, mm = mm)
    }
  }
  dea$pData <- pheno
  return(dea)
}
```

## DEA analysis
```{r}
pnd8.pnd15 <- run_DEA(
  mat = counts.filt, pData = pData,
  groupCol = "Group", group1 = "PND8", group2 = "PND15",
  revGroup = T
)

pnd8.adult <- run_DEA(
  mat = counts.filt, pData = pData,
  groupCol = "Group", group1 = "PND8", group2 = "Adult",
  revGroup = T
)

pnd15.adult <- run_DEA(
  mat = counts.filt, pData = pData,
  groupCol = "Group", group1 = "PND15", group2 = "Adult",
  revGroup = T
)
```

# `cpm` Counts data and pData
```{r}
dge <- DGEList(counts = counts.filt, group = pData[colnames(counts.filt), "Group"])
dge <- calcNormFactors(object = dge, method = "TMM")
cpm <- cpm(dge)

data <- list(cpm = cpm, pData = pData[colnames(cpm), ])

save(data,
  file = "./output/cpm_pData.RData", compress = T,
  compression_level = 3
)
```


# Results
```{r}
dea.list <- list(
  `PND8 vs PND15` = as.DEA(pnd8.pnd15$DEA),
  `PND8 vs Adult` = as.DEA(pnd8.adult$DEA),
  `PND15 vs Adult` = as.DEA(pnd15.adult$DEA)
)

normCounts <- list(
  `PND8 vs PND15` = pnd8.pnd15$voom.norm.counts,
  `PND8 vs Adult` = pnd8.adult$voom.norm.counts,
  `PND15 vs Adult` = pnd15.adult$voom.norm.counts
)

voomEList <- list(
  `PND8 vs PND15` = pnd8.pnd15$EList,
  `PND8 vs Adult` = pnd8.adult$EList,
  `PND15 vs Adult` = pnd15.adult$EList
)

dea.limma <- list(
  `PND8 vs PND15` = pnd8.pnd15$DEA,
  `PND8 vs Adult` = pnd8.adult$DEA,
  `PND15 vs Adult` = pnd15.adult$DEA
)
```


## Save RData files
```{r}
save(dea.list,
  file = "./output/dea_SC_Controls.DEA.RData", compress = T,
  compression_level = 3
)

save(normCounts,
  file = "./output/normCounts_voom_SC_Controls.RData", compress = T,
  compression_level = 3
)

save(voomEList,
  file = "./output/voom_EList_SC_Controls.RData", compress = T,
  compression_level = 3
)

save(dea.limma,
  file = "./output/limma_SC_Controls.RData", compress = T,
  compression_level = 3
)
```


# SessionInfo
```{r}
devtools::session_info()
```